---
title: 'Assignment #1'
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)         # for graphing and data cleaning
library(tidymodels)        # for modeling
library(naniar)            # for analyzing missing values
library(vip)               # for variable importance plots
theme_set(theme_minimal()) # Lisa's favorite theme
```

```{r data}
hotels <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-11/hotels.csv')
```


When you finish the assignment, remove the `#` from the options chunk at the top, so that messages and warnings aren't printed. If you are getting errors in your code, add `error = TRUE` so that the file knits. I would recommend not removing the `#` until you are completely finished.

## Setting up Git and GitHub in RStudio

Read the [Quick Intro](https://advanced-ds-in-r.netlify.app/posts/2021-01-28-gitgithub/#quick-intro) section of the Using git and GitHub in R Studio set of Course Materials. Set up Git and GitHub and create a GitHub repo and associated R Project (done for you when you clone the repo) for this homework assignment. Put this file into the project. You should always open the R Project (.Rproj) file when you work with any of the files in the project. 

**Task**: Below, post a link to your GitHub repository.

https://github.com/neelloc28/Advanced-datasci-hw1.git


## Creating a website

You'll be using RStudio to create a personal website to showcase your work from this class! Start by watching the [Sharing on Short Notice](https://rstudio.com/resources/webinars/sharing-on-short-notice-how-to-get-your-materials-online-with-r-markdown/) webinar by Alison Hill and Desirée De Leon of RStudio. This should help you choose the type of website you'd like to create. 

Once you've chosen that, you might want to look through some of the other *Building a website* resources I posted on the [resources page](https://advanced-ds-in-r.netlify.app/resources.html) of our course website. I highly recommend making a nice landing page where you give a brief introduction of yourself. 


**Tasks**:

* Include a link to your website below. (If anyone does not want to post a website publicly, please talk to me and we will find a different solution).

[My website](https://colleenminnihan.netlify.app)

* Listen to at least the first 20 minutes of "Building a Career in Data Science, Chapter 4: Building a Portfolio". Go to the main [podcast website](https://podcast.bestbook.cool/) and navigate to a podcast provider that works for you to find that specific episode. Write 2-3 sentences reflecting on what they discussed and why creating a website might be helpful for you.

**Jacqueline and Emily mentioned how creating a portfolio (website) can serve multiple purposes, from motivating you to gain new skills, to getting you involved in the data science community, to helping you have something to show at a job interview.  I felt inspired to potentially pursue personal projects in the future when they suggested that you think of a question that genuinely interests you and take it from there (even if it doesn't specifically have to do with the industry you are in/hope to be in), because you can apply those skills you learn in other contexts. I think that creating a website will be helpful for me so that I can have a consolidated space that holds my projects, which will hopefully inspire me to continuously add to it over time.**

* (Optional) Create an R package with your own customized `gpplot2` theme! Write a post on your website about why you made the choices you did for the theme. See the *Building an R package* and *Custom `ggplot2` themes* [resources](https://advanced-ds-in-r.netlify.app/resources.html). 

## Machine Learning review and intro to `tidymodels`

Read through and follow along with the [Machine Learning review with an intro to the `tidymodels` package](https://advanced-ds-in-r.netlify.app/posts/2021-03-16-ml-review/) posted on the Course Materials page. 

**Tasks**:

1. Read about the hotel booking data, `hotels`, on the [Tidy Tuesday page](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-02-11/readme.md) it came from. There is also a link to an article from the original authors. The outcome we will be predicting is called `is_canceled`. 

  - Without doing any analysis, what are some variables you think might be predictive and why?  
  
  **I think that lead_time (days elapsed between entering the date of the booking into the PMS and the arrival date) might be predictive because those that are booked farther out ahead of time may be more likely to be cancelled (due to things coming up closer to the date of stay that would cause it to be cancelled).  I think previous_cancellations could be a predictor as well, because that may be a proxy for how reliable that customer is with committing to their booking.  I also think that children (number of children) could be a predictor because if there are no children (and if it's only one adult), then it may be more likely a business trip which isn't as flexible as, say, a family vacation.**
  
  _ What are some problems that might exist with the data? You might think about how it was collected and who did the collecting.  
  
  **Since customers often change the number of people, length of stay, and preferred room type once they arrive to the hotel, the data may not capture the true distribution of these variables (since the dataset only represents a snapshot in time).  It also may be problematic that the data was only collected during summer months, when people are more likely to be on vacation (which could change their probability of cancelling versus if it was for work).**
  
  - If we construct a model, what type of conclusions will be able to draw from it?  
  
  **If we construct a model, hopefully we will be able to tell what factors (about the person booking, as well as the hotel) may make the booking more likely to be cancelled.  I can see how this could be problematic if the model is put into use because it could potentially result in discrimination for who hotels allow to book (if the hotel deems them likely to cancel).**
  
2. Create some exploratory plots or table summaries of the data, concentrating most on relationships with the response variable. Keep in mind the response variable is numeric, 0 or 1. You may want to make it categorical (you also may not). Be sure to also examine missing values or other interesting values.

```{r, eval = FALSE}
#reading in the data
hotels <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-11/hotels.csv')

#checking to see the distribution of cancelled vs. not cancelled reservations
ggplot(hotels, aes(x = is_canceled)) +
  geom_bar()

#seeing if lead time impacts cancellation
ggplot(hotels, aes(x= lead_time, y = is_canceled)) +
  geom_bar(stat = 'identity') +
  ggtitle("Days booked in advance vs. reservation cancelled")
#hmm this doesn't help me a ton because maybe just more people book less in advance to begin with... Perhaps making this into a categorical variable (less than a week, a week to a month, etc.) would be more enlightening.

ggplot(hotels, aes(x= children, y = is_canceled)) +
  geom_bar(stat = 'identity')
#perhaps would be more helpful to binarize (children Y/N)

#make a new dataset that calculates the percent of cancellations w/in each subset of number of children
children_perc <- hotels %>%
  group_by(children) %>%
  summarize(perc.cancel = sum(is_canceled==0)/sum(is_canceled==1))

#I notice that there are some NA values and children=10, both of which seem a bit odd to me...
ggplot(hotels, aes(x= children)) +
  geom_histogram()

ggplot(children_perc, aes(x=children, y = perc.cancel)) +
  geom_bar(stat = 'identity') +
  ggtitle("number of children and percent cancellation")

#From this histogram, we can see that an overwhelming number of customers booked with 0 kids

#binarizing previous cancellations
hotels <- hotels %>%
  mutate(previous_cancellations_bin = ifelse(previous_cancellations == 0, "no", ifelse(previous_cancellations > 0, "yes", NA)))

ggplot(hotels, aes(x = previous_cancellations_bin)) +
  geom_bar()
#Also, there were barely any people who had a previous cancellation.
```
Distributions of the quantitative variables
```{r}
hotels %>% 
  select(where(is.numeric)) %>% 
  pivot_longer(cols = everything(),
               names_to = "variable", 
               values_to = "value") %>% 
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(vars(variable), 
             scales = "free")

#one weird spike in _date_day_of_?
#most are very right-skewed or have the majority of values at 0 or 1 (ex. lead_time, booking_change, babies, adults)
```

Check the class of each variable:
```{r}
lapply(hotels, class)
```

None of the variables are categorical (or binary) at the moment, but ideally hotel, is_cancelled, stays variables, adults, children, meals, etc. should be.


Now, let's examine the missing values in this dataset.
```{r}
gg_miss_var(hotels)
#It looks as though children is the only variable with missing values, and there are only four missing values, so I don't think too many NAs is a problem in this dataset

#downsize dataset to a few variables we may care most about
hotels_small <- hotels %>%
  select(hotel, is_canceled, lead_time, children, previous_cancellations, adults)
  
vis_miss(hotels_small)
#looks like there are barely any missing values!

ggplot(hotels_small, aes(x = children, y = as.factor(is_canceled))) +
  geom_miss_point()
#So all of the missing values in the children variable are reservations that were cancelled...
```



3. First, we will do a couple things to get the data ready, including making the outcome a factor (needs to be that way for logistic regression), removing the year variable and some reservation status variables, and removing missing values (not NULLs but true missing values). Split the data into a training and test set, stratifying on the outcome variable, `is_canceled`. Since we have a lot of data, we're going to split the data 50/50 between training and test. I have already `set.seed()` for you. Be sure to use `hotels_mod` in the splitting.

```{r}
hotels_mod <- hotels %>% 
  mutate(is_canceled = as.factor(is_canceled)) %>% 
  mutate(across(where(is.character), as.factor)) %>% 
  select(-arrival_date_year,
         -reservation_status,
         -reservation_status_date) %>% 
  add_n_miss() %>% 
  filter(n_miss_all == 0) %>% 
  select(-n_miss_all)

set.seed(494)
```

Split data into training and testing (50/50)
```{r}
#randomly assign 50% of the data to training
hotels_split <- initial_split(hotels_mod, prop = .5)

hotels_split

#training and testing datasets
hotels_training <- training(hotels_split)
hotels_testing <- testing(hotels_split)
```


4. In this next step, we are going to do the pre-processing. Usually, I won't tell you exactly what to do here, but for your first exercise, I'll tell you the steps. 

* Set up the recipe with `is_canceled` as the outcome and all other variables as predictors (HINT: `~.`).  
* Use a `step_XXX()` function or functions (I think there are other ways to do this, but I found `step_mutate_at()` easiest) to create some indicator variables for the fol`lowing variables: `children`, `babies`, and `previous_cancellations`. So, the new variable should be a 1 if the original is more than 0 and 0 otherwise. Make sure you do this in a way that accounts for values that may be larger than any we see in the dataset.  

#are these supposed to be numeric or categorical?

* For the `agent` and `company` variables, make new indicator variables that are 1 if they have a value of `NULL` and 0 otherwise. 
* Use `fct_lump_n()` to lump together countries that aren't in the top 5 most occurring. 
* If you used new names for some of the new variables you created, then remove any variables that are no longer needed. 
* Use `step_normalize()` to center and scale all the non-categorical predictor variables. (Do this BEFORE creating dummy variables. When I tried to do it after, I ran into an error - I'm still investigating why.)
* Create dummy variables for all factors/categorical predictor variables (make sure you have `-all_outcomes()` in this part!!).  
* Use the `prep()` and `juice()` functions to apply the steps to the training data just to check that everything went as planned.
```{r}
#for some reason doing fct_lump_n in the pipe gives me an error? So I'm just doing it for both datasets here so I don't forget
hotels_training$country = fct_lump_n(hotels_training$country,n=5)
hotels_testing$country = fct_lump_n(hotels_testing$country,n=5)

hotels_recipe <- recipe(is_canceled ~ ., data = hotels_training) %>%
  step_mutate(children = ifelse(children > 0,"1","0"), babies = ifelse(babies > 0,"1","0"), previous_cancellations = ifelse(previous_cancellations > 0,"1","0")) %>%
  step_mutate(agent_null = ifelse(agent == 'NULL', "1", "0"), company_null = ifelse(company=='NULL',"1","0")) %>%
  step_rm(agent, company) %>%
  step_normalize(all_predictors(),-all_nominal()) %>% #numeric variables will now have standard deviation of 1 and mean of zero
  step_dummy(all_nominal(),-all_outcomes())


#using prep and juice to make sure everything went as planned
hotels_recipe %>%
  prep(hotels_training) %>%
  juice()
```


5. In this step we will set up a LASSO model and workflow.

* In general, why would we want to use LASSO instead of regular logistic regression? (HINT: think about what happens to the coefficients).

**We would want to use LASSO instead of regular logistic regression to better avoid overfitting our model to the training dataset (thereby having it do poorly with the testing and other datasets).  LASSO assigns a penalty to less important coefficients to shrink them to zero.  The penalty ($\lambda$ can be from 0 to infinity) that minimizes the sum of the squared residuals is the one that LASSO chooses.**

* Define the model type, set the engine, set the `penalty` argument to `tune()` as a placeholder, and set the mode.  
* Create a workflow with the recipe and model.  

6. In this step, we'll tune the model and fit the model using the best tuning parameter to the entire training dataset.

* Create a 5-fold cross-validation sample. We'll use this later. I have set the seed for you.  
* Use the `grid_regular()` function to create a grid of 10 potential penalty parameters (we're keeping this sort of small because the dataset is pretty large). Use that with the 5-fold cv data to tune the model.  
* Use the `tune_grid()` function to fit the models with different tuning parameters to the different cross-validation sets.  
* Use the `collect_metrics()` function to collect all the metrics from the previous step and create a plot with the accuracy on the y-axis and the penalty term on the x-axis. Put the x-axis on the log scale.  
* Use the `select_best()` function to find the best tuning parameter, fit the model using that tuning parameter to the entire training set (HINT: `finalize_workflow()` and `fit()`), and display the model results using `pull_workflow_fit()` and `tidy()`. Are there some variables with coefficients of 0?

```{r}
set.seed(494) # for reproducibility


```


7. Now that we have a model, let's evaluate it a bit more. All we have looked at so far is the cross-validated accuracy from the previous step. 

* Create a variable importance graph. Which variables show up as the most important? Are you surprised?  
* Use the `last_fit()` function to fit the final model and then apply it to the testing data. Report the metrics from the testing data using the `collet_metrics()` function. How do they compare to the cross-validated metrics?
* Use the `collect_predictions()` function to find the predicted probabilities and classes for the test data. Save this to a new dataset called `preds`. Then, use the `conf_mat()` function from `dials` (part of `tidymodels`) to create a confusion matrix showing the predicted classes vs. the true classes. What is the true positive rate (sensitivity)? What is the true negative rate (specificity)? See this [Wikipedia](https://en.wikipedia.org/wiki/Confusion_matrix) reference if you (like me) tend to forget these definitions.
* Use the `preds` dataset you just created to create a density plot of the predicted probabilities of canceling (the variable is called `.pred_1`), filling by `is_canceled`. Use an `alpha = .5` and `color = NA` in the `geom_density()`. Answer these questions: a. What would this graph look like for a model with an accuracy that was close to 1? b. Our predictions are classified as canceled if their predicted probability of canceling is greater than .5. If we wanted to have a high true positive rate, should we make the cutoff for predicted as canceled higher or lower than .5? c. What happens to the true negative rate if we try to get a higher true positive rate? 

8. Let's say that this model is going to be applied to bookings 14 days in advance of their arrival at each hotel, and someone who works for the hotel will make a phone call to the person who made the booking. During this phone call, they will try to assure that the person will be keeping their reservation or that they will be canceling in which case they can do that now and still have time to fill the room. How should the hotel go about deciding who to call? How could they measure whether it was worth the effort to do the calling? Can you think of another way they might use the model? 

9. How might you go about questioning and evaluating the model in terms of fairness? Are there any questions you would like to ask of the people who collected the data? 






## Bias and Fairness

Listen to Dr. Rachel Thomas's  [Bias and Fairness lecture](https://ethics.fast.ai/videos/?lesson=2). Write a brief paragraph reflecting on it. You might also be interested in reading the [ProPublica article](https://www.propublica.org/article/machine-bias-risk-assessments-in-criminal-sentencing) Dr. Thomas references about using a tool called COMPAS to predict recidivism. Some questions/ideas you might keep in mind:

* Did you hear anything that surprised you?  
* Why is it important that we pay attention to bias and fairness when studying data science?  
* Is there a type of bias Dr. Thomas discussed that was new to you? Can you think about places you have seen these types of biases?


